---
title: Manipulate semantic tree of R language
author: Chenliang Xu
date: 2014-11-09 21:30
template: post.html
tags: R
---

_This is basically a reading note on chapter
[Non-standard evaluation of book The Advanced R](http://adv-r.had.co.nz/Computing-on-the-language.html)._

R provides a few functions that expose and manipulate the semantic
tree of R language. The simplest one is `quote`, which return the
semantic tree as is
``` {r}
quote_for <- quote(for(x in 1:3){ s <- s + x})
quote_for
``` 
It works on semantic level,
so the input must be a valid R expression,
``` {r}
## This will lead to an error
## quote(for(x in 1:3 in 2:4){ s <- s + x})
```
The returned result is a R language object, composed of calls and
symbols. We can get elements from the semantic tree,
``` {r}
quote_for[[1]]; quote_for[[2]]; quote_for[[3]]; quote_for[[4]];
```
and even set elements,

``` {r}
quote_for[[3]] <- quote(1:10)
quote_for
s  <- 0
eval(quote_for)
s
```

The function `substitute(expr, env)` substitute elements in semantic tree. Its
behavior is unnecessarily complex,
- If the argument `env` is `.GlobalEnv`, works the same as `quote`, without any substitute
- If the argument `env` is not `.GlobalEnv`
 - If the argument `expr` is a promise object, get the expression slot
   of the promise without any substitute. The is typically used to
   capture unevaluated expression from caller function.
 - If the argument `expr` is not a promise object, substitute the
    unevaluated semantic tree of `expr` using the bound in `env`

This is a very bad design, that a function might do totally different
things depending on the running environment. Such a complex design
makes it hard to guess its behavior.

Following suggestions in book _Advanced R_, we define a few helper
functions, each of which does only one job clearly. Note the name of
functions are different from Advanced R. The suffix `*_s` means
standard evaluation, so it's the caller's responsibility to quote code
explicitly; `*_q` means the arguments are quoted automatically. Lisp
users may find name `*_q` familiar, similar to `setq`.
``` {r}
## The definition here may be hard to follow. Please focus on the
## usage explained below, instead of the ugly implementation. If you
## are curiouse about the implementation, please read Advaned R.
substitute_q <- function(expr, env = emptyenv()) {
  expr <- substitute(expr)
  if (identical(env, globalenv())) {
    env <- as.list(env)
  }
  eval(substitute(substitute(expr, env), list(expr = expr)))
}

substitute_s <- function(expr, env = emptyenv()) {
  call <- substitute_q(substitute_q(x, env), list(x = expr))
  eval(call)
}
```

`substitute_s` substitute the symbols in `expr` at lexical basis. For example,
``` {r}
quote_for <- quote(for(x in 1:3){ s  <- s + x})
substitute_s(quote_for, list(x = quote(y)))
```
`substitute_q` works similarly, but it quote the argument `expr`
automatically, so the caller doesn't need to quote the code and must
pass the argument `expr` lexically. It seems to be convenient, but may
cause some issue since we cannot pass the quoted code saved in a
variable to `substitute_q`.
``` {r}
substitute_q(for(x in 1:3){ s  <- s + x}, list(x = quote(y)))
```

`bquote` is similar but provides better control. It return unevaluated
semantic tree, and only substitute symbols wrapped in `.()`.
``` {r}
## The result is not meaningful. Why do I do that? Because I can.
bquote(for(x in 1:3){ s  <- s + .(x)}, list(x = quote(y)))
```

Both `substitute_*` and `bquote` work on lexical basis, so there is no
guarantee that the result is semantically correct or
meaningful. Nonetheless, it is much easier and safer to manipulate R
code with those functions than manipulate R code as strings.
